---
title: "attritionEDA"
author: "Ryan Kinney"
date: "2/21/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(skimr)
library(dplyr)
library(ggplot2)
library(stringr)
library(caret)
library(class)
library(magrittr)
library(dlookr)
library(ggthemes)
library(plyr)
library(dplyr)
library(e1071)
```

```{r import}
CaseStudy2data = read.csv("../casestudy02/CaseStudy2data.csv", header = TRUE) #data set
CaseStudy2dataNoAttrition = read.csv("../casestudy02/CaseStudy2dataNoAttrition.csv", header = TRUE) #no attrition
CaseStudy2dataNoSalaries = read.csv("../casestudy02/CaseStudy2dataNoSalary.csv", header = TRUE) #no salary
```

#DDS Analytics specializes in talent management solutions for Fortune 100 companies, with predicting turnover
#as the first application of data science for talent management. In this Case study, we have been given a dataset 
# to identify factors leading to attrition. Then, a model will be created to predict attrition. Afterwards, we 
# will analyze factors correlated with employee salary, and build a model to predict employee salaries.

##Data cleanup

```{r}
#check for character variables and convert to factors:
CategoricalVars <- names(CaseStudy2data %>% select(where(is.character)))
CaseStudy2data[sapply(CaseStudy2data,is.character)] <- lapply(CaseStudy2data[sapply(CaseStudy2data, is.character)], as.factor)

```

## Looking at the summary statistics and Job happiness in each department:

```{r attrition EDA}
summary(CaseStudy2data)
CaseStudy2data %>% group_by(Attrition) %>% skim()
#*non-relevant variables with zero impact on attrition: ID, Employee count, employee number
Attrition.cleaned = CaseStudy2data %>% select(-ID, -EmployeeCount, -EmployeeNumber)
#comparing job ratings between jobs. R&D manufacturing Directors and Human Resources have the highest ratings
Attrition.cleaned %>%
  group_by(Department, JobRole) %>% 
    summarize(Job_Satisfaction=mean(JobSatisfaction),
              Relationship_Satisfaction=mean(RelationshipSatisfaction),
              Performance_Rating = mean(PerformanceRating),
              average_ratings = ((Job_Satisfaction + Relationship_Satisfaction + Performance_Rating)/3))
#Overtime playing a big role in attrition
CaseStudy2data %>% group_by(OverTime,Attrition) %>% count()
 
 
```
#3 Factors Contributing to Turnover:

## Factor 1 : Department
```{r}

Attrition.cleaned %>% group_by(Attrition) %>% count(Department) 
#attrition rates : Human Resources = 17.1%; R&D = 13.3%; Sales = 21.6%
department.attrition.rate = data.frame(Department = c("Human Resources", "Research and Development", "Sales"), 
                                       AttritionRate = c(17.1,13.3,21.6))

#bar graph showing attrition for each department. Sales has the highest attrition rate
department.attrition.rate %>% 
  ggplot(aes(x=Department, y= AttritionRate)) + 
  geom_bar(position= "dodge",stat="identity", fill ="palevioletred4") +
  geom_text(aes(label = AttritionRate, y = AttritionRate),vjust = -1, position = position_dodge(width = .9)) +
  ggtitle("Attrition Rates in each department") +
  theme_minimal() +
  ylab("Attrition Rate (percentage)")
#count of people in each department + attrition
CaseStudy2data %>% 
  ggplot(aes(x= Department, fill=Attrition)) +
  geom_bar(position = "dodge") +
  geom_text(aes(label = ..count..),stat="count", vjust = -.5, position = position_dodge(.9)) +
  ggtitle("Number of Employees in Each Department") +
  ylab("Count") +
  theme_minimal()

```

##Factor 2: Years at Company
```{r}
Attrition.cleaned%>% group_by(Attrition) %>% summarise(mean(YearsAtCompany))

#boxplot showing the attrition explained by age
CaseStudy2data %>% ggplot(aes(x=Attrition, y=YearsAtCompany)) + 
  geom_boxplot() +
  scale_fill_grey() +
  ggtitle("Attrition Explained By Years at Company") +
  theme_economist() +
  ylab("Years at the Company")


```

##Factor 3: Marital Status
```{r}
Attrition.cleaned %>% group_by(Attrition) %>% count(MaritalStatus)
  #attrition rates: Divorced = 6.28%; married = 14.5%; single = 26.02%
marital.attrition.rate = data.frame(MaritalStatus = c("Divorced", "Married", "Single"), 
                                    AttritionRate = c(6.28,14.15,26.02))

#bar chart showing attrition rate of each marital status                                   
marital.attrition.rate %>% 
  ggplot(aes(x=MaritalStatus, y= AttritionRate)) + 
  geom_bar(position= "dodge",stat="identity", fill = "palevioletred4") +
  geom_text(aes(label = AttritionRate, y = AttritionRate),vjust = -.3, position = position_dodge(width = .9), size = 5) +
  ggtitle("Attrition Rates of each Marital Status") +
  theme_minimal() +
  xlab("Marital Status") +
  ylab("Attrition Rate (percentage)")

#A look at marital status among employees
CaseStudy2data %>% 
  ggplot(aes(x= MaritalStatus, fill=Attrition)) +
  geom_bar(position = "dodge") +
  geom_text(aes(label = ..count..),stat="count", vjust = -.5, position = position_dodge(.9)) +
  ggtitle("Number of Employees Who Are Divorced, Married, or Single") +
  xlab("Marital Status") +
  ylab("Count") +
  theme_minimal()

#density plot for monthly income based on marital status
CaseStudy2data %>% 
  ggplot(aes(x=MonthlyIncome)) + 
  geom_density(aes(fill=MaritalStatus), alpha=.6) + 
  ggtitle("Density Plot of Monthly Income by Marital Status") +
  xlab("Monthly Income")
```



#Final KNN model with 81.99% accuracy rating, 82.87% Sensitivity, and 60% specificity rating

```{r}

#KNN
set.seed(11)
splitPerc = .7
trainIndices = sample(1:dim(CaseStudy2data)[1], round(splitPerc*dim(CaseStudy2data)[1]))
train = CaseStudy2data[trainIndices,]
test = CaseStudy2data[-trainIndices,]

#best K-value:
accs = data.frame(accuracy = numeric(30), k = numeric(30))
for(i in 1:30)
{
  classifications = knn(train[,c(2,5,20,32,33)],test[,c(2,5,20,32,33)],train$Attrition,prob=TRUE, k=i)
  table(test$Attrition,classifications)
  CM = confusionMatrix(table(test$Attrition,classifications))
  accs$accuracy[i] = CM$overall[1]
  accs$k[i] = i
}
plot(accs$k,accs$accuracy, type = "l", xlab = "K") #15 best value for k but lowers specificity to 50%

#build model and look at confusion matrix statistics
classifications = knn(train = train[,c(2,5,20,32,33)],test = test[,c(2,5,20,32,33)],cl = train$Attrition, prob = TRUE, k =7)
confusionMatrix(table(test$Attrition,classifications))

```


#Final Naive Bayes model adding factor variables. Accuracy = 80.46%, Sensitivity = 97.17%, Specificity = 8.16%

```{r}
#Naive Bayes
model = naiveBayes(Attrition ~ Age + Department + Gender + HourlyRate + JobLevel + JobSatisfaction + MaritalStatus, data = train)
CM = confusionMatrix(table(predict(model,test),test$Attrition))

```


#Predicting using a data set without attrition

```{r}
classifications = knn(train = train[,c(2,5,20,32,33)],test = CaseStudy2dataNoAttrition[,c(2,4,19,31,32)], cl = train$Attrition, prob = TRUE, k= 7)

KNNpredictions = as.data.frame(classifications)
KNNpredictions$ID <- CaseStudy2dataNoAttrition %>% select(ID)
write.csv(KNNpredictions, "../casestudy02/CaseStudy2KNNpredictions.csv", quote = FALSE)

```


#Creating regression model to be used to test the no salary data set

```{r}
#Variables that have highest R-sq value in relation to Monthly Income :
fit1 = lm(MonthlyIncome~JobLevel, data = CaseStudy2data)
summary(fit1) #r-sq = .9055
fit2 = lm(MonthlyIncome~StockOptionLevel, data = CaseStudy2data)
summary(fit2) #not good
fit3 = lm(MonthlyIncome~DistanceFromHome, data = CaseStudy2data)
summary(fit3) #not good
fit4 = lm(MonthlyIncome~JobRole, data = CaseStudy2data)
summary(fit4) #r-sq = .8087
fit5 = lm(MonthlyIncome~TotalWorkingYears, data = CaseStudy2data)
summary(fit5) #r-sq = .6056
fit6 = lm(MonthlyIncome~Education, data = CaseStudy2data)
summary(fit6) #not good
fit7 = lm(MonthlyIncome~BusinessTravel, data = CaseStudy2data)
summary(fit7) #not good
fit8 = lm(MonthlyIncome~YearsAtCompany, data = CaseStudy2data)
summary(fit8) #r-sq = .24
fit9 = lm(MonthlyIncome~YearsInCurrentRole, data = CaseStudy2data)
summary(fit9) #r-sq = .13
fit10 = lm(MonthlyIncome~YearsSinceLastPromotion, data = CaseStudy2data)
summary(fit10) #not good

#combined model
modelfit = lm(MonthlyIncome~JobLevel + JobRole+ TotalWorkingYears + YearsAtCompany + YearsInCurrentRole, data = CaseStudy2data)
summary(modelfit) #r-sq = .9472
confint(fit10)
rmse <- sqrt(mean(residuals(modelfit)^2)) #1055.69 < 3000

#Salary Predictions using modelfit

Salaries = predict(modelfit, newdata = CaseStudy2dataNoSalaries)
SalaryPredictions = as.data.frame(Salaries)
SalaryPredictions$ID <- CaseStudy2dataNoSalaries %>% select(ID)
write.csv(SalaryPredictions, "../casestudy02/CaseStudy2Salarypredictions.csv", quote = FALSE)



```


# youtube link : https://www.youtube.com/watch?v=wvL-XCkBzgo
